---
title: "Regressions"
subtitle: "Correlations for the Advanced?"
author: "Erik Kusch"
date: "01/04/2020"
fontsize: 10pt
output: 
  beamer_presentation: 
    keep_tex: true # keep latex file that is generated during compilation
    toc: false # this is added through a later command
    slide_level: 3 # at how many pound signs (#) to assume slide title level
    includes:
      in_header: Style.tex
classoption: c # change to `handout` to ignore breaks
---
  
```{r setup, include=FALSE}
options(scipen=1, digits=4)
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, cache = TRUE, cache.lazy=TRUE, tidy.opts=list(width.cutoff=50),tidy=FALSE, fig.height=8, digits=4)
```

<!--- ####### TOC (use this if you want ## headers to be included in the toc)--------------------
---------------------------------------  --->
  
\tableofcontents

<!--- ####### DOCUMENT --------------------
---------------------------------------  --->

# The Basics
## Correlation Tests
### Terminology
\begin{center}
Correlation is \textbf{not} necessarily \textbf{causation} (\raggedleft \href{https://tylervigen.com/page?page=1}{spurious correlations}).
\end{center}

\begincols[T]
\begincol{.5\linewidth}
  \pause
  Correlation tests yield two measurements:
  \begin{itemize}
  \item \textit{r} value (measure of correlation)
    \begin{itemize}
    \item $r \approx 1$ (strong, positive correlation)
    \item $r \approx 0$ (no correlation)
    \item $r \approx -1$ (strong, negative correlation)
    \end{itemize}
  \item \textit{p} value (measure of statistical significance)
  \end{itemize}
\endcol
\begincol{.5\linewidth}
\vspace{30pt}
  \centering \includegraphics[width=.9\linewidth]{Figures/coeffs.png}
\endcol
\endcols
\pause
\vspace{.5cm}
\raggedleft $\rightarrow$ Get a feeling for it on \href{http://guessthecorrelation.com/}{Guess The Correlation}.

### Types of Correlations
\begin{center}
These approaches are extremely useful in data exploration and for preliminary analyses!
\end{center}

\begincols[T]
\begincol{.5\linewidth}
  \pause
  Prominent correlation tests include:
  \begin{itemize}
  \item Contingency Coefficient
  \item Kendall's Tau
  \item Spearman Correlation
  \item Pearson Correlation
  \item Cramer's V
  \item ANalysis Of VAriance (ANOVA)
  \item ...
  \end{itemize}
\endcol
\begincol{.5\linewidth}
  \centering \includegraphics[width=.9\linewidth]{Figures/Correl.png}
\endcol
\endcols

## Regression Models
### Terminology
\begincols
\begincol{.5\linewidth}
  \pause
  \begin{itemize}
    \item $\alpha$ - The \textbf{Intercept}. The value of $y$ when $x = 0$ (also referred to as $\beta_0$).
    \item $\beta_i$ - The \textbf{Correlation Coefficient}. The increase in $y$ for a one-unit increase in dependent variable $i$ (usually, $x$ if only one dependent variable).
    \item $\epsilon$ - The \textbf{Random Error}. The deviation of data points from the regression line. Usually assumed to follow $\epsilon \sim N(0, \sigma^2)$
  \end{itemize}
\endcol
\begincol{.5\linewidth}
  \centering \includegraphics[width=1\linewidth]{Figures/Regr.png}
  \raggedleft \tiny Modified after Knut Helge Jensen.
\endcol
\endcols

### Assumptions in Theory
Linear regression models need to be inspected for violations of assumptions after regressing:
  \begin{itemize}
    \item \textbf{Residuals vs. Fitted values} \linebreak 
    Non-linear patterns identify a non-linear relationship between dependent and independent variables.
    \item \textbf{Normal Q-Q plot} \linebreak
    Non-normal distribution of residuals shows that the assumption of $\epsilon \sim N(0, \sigma^2)$ is violated.
    \item \textbf{Scale Location} \linebreak
    Non-constant variance identifies show that the assumption of homoscedasticity (invoked by least squares fitting).
    \item \textbf{Residuals vs. Leverage} \linebreak
    A non-zero trend identifies the presence of influential outliers.
  \end{itemize}
  
### Assumptions in `R`
\begincols
\begincol{.45\linewidth}
  \pause
  \begin{itemize}
    \item Simply type `plot(...)` with `...` denoting your regression model.
    \item You can also target individual plots by writing:
    \begin{itemize}
      \item `plot(..., 1)` for Residuals vs. Fitted values
      \item `plot(..., 2)` for Normal Q-Q plot
      \item `plot(..., 3)` for Scale Location
      \item `plot(..., 4)` for Residuals vs. Leverage
    \end{itemize}
  \end{itemize}
\endcol
\begincol{.55\linewidth}
  \centering \includegraphics[width=1\linewidth]{Figures/Assumpts.png}
  \raggedleft \tiny By Knut Helge Jensen.
\endcol
\endcols



### Types of Regressions
\begin{center}
Less model variables result in a more interpretable model!
\end{center}

\vspace{1cm}

\begincols[T]
\begincol{.5\linewidth}
  Prominent regression approaches include the following:
  \pause
  \begin{itemize}
  \item \textbf{Single Linear Regression}
  \item Multiple Linear Regression
  \item \textbf{Linear Mixed Effect Models}
  \item \textbf{Generalized Linear Models}
  \end{itemize}
\endcol
\begincol{.5\linewidth}
  \pause
  \begin{itemize}
  \item Polynomial Regressions
  \item Generalized Additive Models
  \item Regression Splines
  \item Smoothing Splines
  \item Local Regressions
  \item ...
  \end{itemize}
\endcol
\endcols


## Least Squares vs. Maximum Likelihood
### Least Squares vs. Maximum Likelihood
\begin{center}
These methods refer to \textbf{parameter estimation}.
\end{center}

\vspace{.3cm}

\begincols[T]
\begincol{.5\linewidth}
  \textit{Ordinary Least Squares} (OSL):
  \begin{itemize}
  \item Used for most basic linear regressions
  \item obtain coefficient estimates $\hat{\beta}_{0}$ and $\hat{\beta}_{1}$ such that the linear model fits the available data well — that is, so that $\hat{y}_{i} \approx \hat{\beta}_{0} + \hat{\beta}_{1}x_{i}$ for $i = 1, ..., n$.
  \end{itemize}
\endcol
\begincol{.5\linewidth}
  \textit{Maximum Likelihood Estimation} (MLE):
  \begin{itemize}
  \item Used in logistic regressions and generalized linear models
  \item estimates for $\beta_{0}$ and $\beta_{1}$ such that the predicted probability $\hat{Pr}(x_{j})$ corresponds to the observed response variable status.
  \end{itemize}
\endcol
\endcols

\vspace{.5cm}

\begincols[T]
\begincol{.5\linewidth}
  Minimize:  
  \tiny
  \begin{equation}
  RSS = \sum\limits_{i=1}^n(y_{i} -\hat{y}_{i})^2
  \end{equation}
\endcol
\begincol{.5\linewidth}
  Maximize:  
  \tiny
  \begin{equation}
  \ell(\theta) = \prod\limits_{i=1}^{n}f(x_i|\theta)
  \end{equation}
\endcol
\endcols
with $\hat{y}_{i} = \hat{\beta}_{0} + \hat{\beta}_{1}x_{i}$
\normalsize

# Methods \& Models
## Single Linear Regression
### Purpose & Assumptions
\textbf{Single linear regression} \flushright `lm()` in base `R`  
\vspace{-5pt}
\begin{table}[c]
  \centering
  \resizebox{\textwidth}{!}{%
  \begin{tabular}{L{0.2\linewidth}L{0.8\linewidth}}
    \textit{Purpose:} &  Identify whether and how two variables are related.\\
    \pause
    \textit{Assumptions:} & 
      \begin{itemize}
      \item Down to \textit{Study-Design}:
      \begin{itemize}
      \item Predictor variable is continuous (ratio or interval scale)
      \item Response variable is continuous (ratio or interval scale)
      \item Variable values are \textbf{independent} (not paired)
      \end{itemize}
      \item Need for \textit{Post-Hoc Tests}:
      \begin{itemize}
      \item Variable values follow \textbf{homoscedasticity} (equal variance across entire data range)
      \item Residuals follow normal distribution (\textbf{normality})
      \item Absence of \textbf{influential outliers}
      \item Response and Predictor are related in a \textbf{linear} fashion
      \end{itemize}
      \end{itemize}
     \\
  \end{tabular}}
\end{table}


### Example - The Data
\footnotesize
```{r SingleLin1}
# measures of Diameter (labelled as Girth), Height, and Volume of Timber
data("trees")
head(trees) 
```
\vspace{.5cm}
\raggedleft $\rightarrow$ Let's see if there is a good regression to be had between *Girth* and *Volume*.

### Example - The Model 
\footnotesize
```{r SingleLin2, fig.height= 5}
SingleLin_Mod <- with(trees, lm(Girth ~ Volume))
par(mfrow=c(2,2))
plot(SingleLin_Mod)
```

### Example - Refining The Model 
\footnotesize
```{r SingleLin3, fig.height= 5}
trees <- trees[-31,] # removing the influential otulier in row 31
SingleLin_Mod <- with(trees, lm(Girth ~ Volume))
par(mfrow=c(2,2))
plot(SingleLin_Mod)
```

### Example - Model Output
\tiny
```{r SingleLin4, fig.height= 5}
summary(SingleLin_Mod)
```
\vspace{.5cm}
At a `Volume` of $0$, `Girth` is predicted to be `r summary(SingleLin_Mod)[["coefficients"]][1,1]` (of course that doesn't make sense, not only is a volume of $0$ biological nonsense, height also plays a part here for sure). For a one-unit increase in `Volume`, `Girth` is predicted to go up by `r summary(SingleLin_Mod)[["coefficients"]][2,1]` inches (yes, they recorded in inches). Both estimates are statistically significant.

## Mixed Effect Models
### Purpose & Assumptions
\textbf{Linear mixed effect model} \flushright `lme()` in base `nlme` package  
\vspace{-5pt}
\begin{table}[c]
  \centering
  \resizebox{\textwidth}{!}{%
  \begin{tabular}{L{0.2\linewidth}L{0.8\linewidth}}
    \textit{Purpose:} &  Identify whether and how variables are related.\\
    \pause
    \textit{Assumptions:} & 
      \begin{itemize}
      \item Down to \textit{Study-Design}:
      \begin{itemize}
      \item Predictor variable is continuous (ratio or interval scale)
      \item Response variables are continuous (ratio or interval scale) and/or categorical (metric or ordinal scale)
      \end{itemize}
      \item Need for \textit{Post-Hoc Tests}:
      \begin{itemize}
      \item Variable values follow \textbf{homoscedasticity} (equal variance across entire data range)
      \item Residuals follow normal distribution (\textbf{normality})
      \item Absence of \textbf{influential outliers}
      \item Response and Predictor are related in a \textbf{linear} fashion
      \end{itemize}
      \end{itemize}
     \\
  \end{tabular}}
\end{table}

### Fixed vs. Random Effects
\begin{center}
Fixed effects and random effects are also referred to as fixed effect factors and random effect factors.
\end{center}

\begincols[T]
\begincol{.5\linewidth}
  \pause
  **Fixed Effects**:
  \pause
  \begin{itemize}
  \item Informative factor levels regarding hypothesis. 
  \item Want to study these levels and their effects.
  \item Factor levels are deliberate part of the study-design.
  \item Higher sample size $\neq$ higher number of levels.
  \end{itemize}
\endcol
\begincol{.5\linewidth}
  \pause
  **Random Effects**:
  \pause
  \begin{itemize}
  \item Uninformative factor levels regarding hypothesis. 
  \item Do not want to study these levels and their effects.
  \item Factor levels are imposed by nature/type of study.
  \item Usually: higher sample size $=$ higher number of levels.
  \end{itemize}
\endcol
\endcols

\vspace{.3cm}
\pause
\raggedleft Usually stored in `R` in `factor` mode/class.

### Example - The Data
\footnotesize
```{r MultiLin1}
# measures of Weight, Diet, Time, and Chicks
data("ChickWeight")
head(ChickWeight) 
```
\vspace{.5cm}
\raggedleft $\rightarrow$ Let's see if there is a good regression to be had between *weight* and *Time* while accounting for random effects belonging to *Chick*, and fixed effects of *Diet*.

### Example - The Model 
\tiny
```{r MultiLin2a, fig.height= 5}
library(nlme)
MultiLin_Base <- lme(weight ~ Time*Diet, # weight as an interaction of time and diet
                     random = ~+1|Chick, # random effect of Chick
                     data = ChickWeight)
```
We now have our model. However, we know that time is a component and we likely have repeated samples here. In these cases, we need to account for auto-correlation by defining a correlation structure.
```{r MultiLin2b, fig.height= 5}
MultiLin_Mod <- lme(weight ~ Time*Diet, random = ~+1|Chick, 
                    cor=corAR1(), # adding autocorrelation structure since we have repeated measures
                    data = ChickWeight)
```
Let's see which model (basic or the one with auto-correlative structure) performs better:
```{r MultiLin2c, fig.height= 5}
anova(MultiLin_Base, MultiLin_Mod) # second model is better
```
We clearly prefer the more sophisticated, auto-correlative model and want to see which of its parameters are informative:
```{r MultiLin2d, fig.height= 5}
anova(MultiLin_Mod) # all parameters should be kept
```
We keep all parameters. Although the inclusion of `Diet` is not significant, the interaction of `Diet` and `Time` is, therefore, both `Time` and `Diet` need to stay irrespective of their significance.

### Example - Assessing the Model
\tiny
```{r MultiLin3, fig.height= 5}
par(mfrow=c(2,2))
plot(fitted(MultiLin_Mod), resid(MultiLin_Mod)) # values around 0 -> good
plot(fitted(MultiLin_Mod), ChickWeight$weight) # pattern fuzzy, but linear -> good
qqnorm(resid(MultiLin_Mod)) # residuals are not normal dsitributed -> bad
```

### Example - Model Output
\tiny
```{r MultiLin4, fig.height= 5}
summary(MultiLin_Mod)
```

### Example - Model Output Explained
Variance between chicks ($0.006581$) than residual variance ($42.46$). This is good!  
\vspace{.2cm}

A chick is expected to have a `weight`of `r summary(MultiLin_Mod)[["coefficients"]][["fixed"]][1]` at `Time` $= 0$ and `Diet` $= 1$. Per time-step, `weight` is expected to increase by `r summary(MultiLin_Mod)[["coefficients"]][["fixed"]][2]`.
\vspace{.2cm}

Mean chick `weight` is different to (read: "Diet1 weights are smaller by") `Diet` $=1$ by `r summary(MultiLin_Mod)[["coefficients"]][["fixed"]][3]`, `r summary(MultiLin_Mod)[["coefficients"]][["fixed"]][4]`, and `r summary(MultiLin_Mod)[["coefficients"]][["fixed"]][5]` for `Diet` $=2$, `Diet` $=3$, and `Diet` $=4$ respectively (take note that these differences aren't statistically significant).  
\vspace{.2cm}

`weight` of chicks increases, on average, increases by `r summary(MultiLin_Mod)[["coefficients"]][["fixed"]][6]` units more per one-unit increase in time when compared to `Diet` $=1$. The same logic applies to `Diet` $=3$, and `Diet` $=4$.

### Example - Model Output Visualised
\tiny
```{r Multi5, fig.height = 4.8}
library(ggplot2)
ChickWeight$fit <- predict(MultiLin_Mod, level=0)
ggplot(ChickWeight, aes(Time, weight)) + 
  geom_jitter(aes(colour=Diet, shape=Diet), width=0.1, size=3) + 
  theme_bw(base_size=20) + 
  geom_line(aes(y=fit, lty=Diet))
```

## Generalised Linear Models
### Purpose & Assumptions
\textbf{Generalized Linear Models} \flushright `glm()` in base `R`  
\vspace{-5pt}
\begin{table}[c]
  \centering
  \resizebox{\textwidth}{!}{%
  \begin{tabular}{L{0.2\linewidth}L{0.8\linewidth}}
    \textit{Purpose:} &  Identify whether and how variables are related.\\
    \pause
    \textit{Assumptions:} & 
      \begin{itemize}
      \item Down to \textit{Study-Design}:
      \begin{itemize}
      \item Variable values are \textbf{independent} (not paired)
      \end{itemize}
      \item Need for \textit{Post-Hoc Tests}:
      \begin{itemize}
      \item Absence of \textbf{influential outliers}
      \item Response and Predictor are related in a \textbf{linear} fashion
      \end{itemize}
      \end{itemize}
     \\
  \end{tabular}}
\end{table}

$\rightarrow$ Allow for \textbf{non-normal} distributions and \textbf{heteroscedasticity}

### Linear Predictor, Link Function, and Variance Function
**Components** of a Generalized Linear Model:
\begin{enumerate}
\item \textbf{Linear predictor} e.g.: $y_i = \alpha + \beta_1 x_i$
\item \textbf{Link function} $g(\hat{y_i}) = y_i$ \linebreak 
Relationship between predictor value and estimated value.
\item \textbf{Variance function} $var(y_i)=\phi V_i(\overline{x})$ \linebreak
Variance depends on predictor mean, dispersion parameter $\phi$ is a constant
\end{enumerate}

\pause 
\vspace{.3cm}

Which **combinations** of components do I use?
\begin{table}[htbp]
  \centering
    \begin{tabular}{rrrr}
    \toprule
    Error & Link function & Variance function &  Typical type of data \\
    \midrule
    normal  & identity  & $1$ (constant) & Textbook examples \\
    Poisson & $log$   & $var=\overline{x}$ & Count data \\
    binomial & logit, $log(\overline{x}/(1-\overline{x}))$ & $var=\overline{x}(1-\overline{x})/n$ & Binary data \\
    \bottomrule
    \end{tabular}
\end{table}

### Example - The Data
\footnotesize
```{r GeneralLin1}
data("ChickWeight")
head(ChickWeight)
```
\vspace{.5cm}
\raggedleft $\rightarrow$ Let's reassess our earlier analysis of chick **weight** as a function of **time** and **diet**. This time, we will forego the random effect of chicks since that would create a generalized linear mixed effect model.

### Example - The Model 
\tiny
```{r GeneralLin2, fig.height= 5}
GeneralLin_Mod <- glm(weight ~ Time*Diet, family = poisson, data = ChickWeight)
par(mfrow=c(2,2))
plot(GeneralLin_Mod)
```

### Example - Model Output
\tiny
```{r GeneralLin3, fig.height= 5}
summary(GeneralLin_Mod)
```
\vspace{.2cm}
As you can see, the estimates to earlier have changed drastically. 

### Example - Model Output Explained
A chick is now expected to have a `weight`of `r summary(GeneralLin_Mod)[["coefficients"]][1,1]` at `Time` $= 0$ and `Diet` $= 1$. Per time-step, `weight` is expected to increase by `r summary(GeneralLin_Mod)[["coefficients"]][2,1]`.
\vspace{.2cm}

Mean chick `weight` of `Diet` $=1$ is smaller by `r summary(GeneralLin_Mod)[["coefficients"]][3,1]`, `r summary(GeneralLin_Mod)[["coefficients"]][4,1]`, and `r summary(GeneralLin_Mod)[["coefficients"]][5,1]` for `Diet` $=2$, `Diet` $=3$, and `Diet` $=4$ respectively (these differences are now significant).  
\vspace{.2cm}

`weight` of chicks increases, on average, increases by `r summary(GeneralLin_Mod)[["coefficients"]][6,1]` units more per one-unit increase in time when compared to `Diet` $=1$. The same logic applies to `Diet` $=3$, and `Diet` $=4$.

# Choosing the Right Method
### Choices, Choices, Choices...

- \textbf{Linear Model} `lm`. When all \textbf{assumptions are met} (i.e.: homoscedasticity, normality, independence).

\vspace{.2cm}

- \textbf{Linear Mixed Effect Model} `lme`. When the assumption of \textbf{independence is violated}.  

\vspace{.2cm}

- \textbf{Generalized Linear Model} `glm`. When the assumptions of \textbf{homoscedasticity} and \textbf{normality} are \textbf{violated}.  

\vspace{.2cm}

- \textbf{Generalized Linear Mixed Effect Model} `glmmPQL` from `MASS`, or `glmer` from `lme4`. When the assumptions of \textbf{homoscedasticity}, \textbf{normality}, and \textbf{independence} are \textbf{violated}.  

