---
title: "Data Handling and Assumptions"
subtitle: "Making the Most of Your Data"
author: "Erik Kusch"
date: "25/03/2020"
fontsize: 10pt
output: 
  beamer_presentation: 
    keep_tex: true # keep latex file that is generated during compilation
    toc: false # this is added through a later command
    slide_level: 3 # at how many pound signs (#) to assume slide title level
    includes:
      in_header: Style.tex
classoption: c # change to `handout` to ignore breaks
---
  
```{r setup, include=FALSE}
options(scipen=1, digits=4)
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, cache = TRUE, cache.lazy=TRUE, tidy.opts=list(width.cutoff=50),tidy=TRUE, fig.height=8, digits=4)
```

<!--- ####### TOC (use this if you want ## headers to be included in the toc)--------------------
---------------------------------------  --->
  
\tableofcontents

<!--- ####### DOCUMENT --------------------
---------------------------------------  --->

# Data Etiquettes
### Why Care?
\begin{tcolorbox}[colback=burgundy!5,colframe=burgundy!40!black,title=]
\centering \textit{Biostatisticians often use 70\% of their time to handle data and just 30\% to actually analyse it.}
\end{tcolorbox}

\begincols[T]
  \begincol{.5\linewidth}
  \pause
    \textbf{Why care?} 
    \begin{itemize}
    \item Proper data collection and data handling ensure accurate results
    \item Proper data collection cuts down on data handling time
    \item Proper data handling will make reproducing an analysis much easier
    \end{itemize}
  \endcol
  \begincol{.5\linewidth}
  \pause
  \textbf{What to consider?} 
    \begin{itemize}
    \item Which data format to use
    \item What kind of data to record
    \item How data values are recorded/stored
    \item What kind of data values are feasible
    \end{itemize}
  \endcol
\endcols

## Data Recording
### Data Recording
\textbf{Guidelines for data recording:}
\begin{itemize}
\pause
  \item When collecting categorical data, know what values the variables are allowed to take
\pause
  \item When collecting continuous data, know which range the variable values can fall into
\pause
  \item Make sure everyone involved in data collection is on the same page
\pause
  \item Make regular back-ups of your data set
\end{itemize}

\vspace{.5cm}

\pause
\textbf{I recommend:}
\begin{itemize}
\pause
  \item Preparing content-aware excel files for data entry
    \begin{itemize}
    \pause
    \item Only allow pre-defined values to be entered
    \pause
    \item Need some excel macro writing
    \end{itemize}
\pause
  \item Using a cloud-service featuring version control for data storage
\end{itemize}

### Common Issues
\pause
\textbf{The Decimals}  
\pause
Always use a \textit{dot} to indicate decimals.  
$\rightarrow$ It is the standard in science.
  
\pause
\vspace{.5cm}
\textbf{To} `NA` \textbf{Or Not To} `NA`\textbf{?}  
\pause
\textit{Never enter} `NA` \textit{values manually} into your data.  
$\rightarrow$ They cause problems in `R`.
  
\pause
\vspace{.5cm}
\textbf{Redundancy Or Sparsity?}  
\pause
\textit{Don't clutter data} with unnecessary data records.  
$\rightarrow$ Reduces storage space and chances for errors.

## Data Storing
### Data Storing
`R` works very well with:  

\pause
\vspace{.1cm}
\begin{itemize}
\item excel files (\textit{.xls, .xlsx, .csv})
\end{itemize}
$\rightarrow$ Easiest to handle outside of `R`, most storage-heavy  
$\rightarrow$ Make sure to provide co-workers with a master file before data collection to avoid cell formatting issues on different computers

\pause
\vspace{.1cm}
\begin{itemize}
\item text files (\textit{.txt})
\end{itemize}
$\rightarrow$ Difficult to handle outside of `R`, easy on storage  
$\rightarrow$ I advise against using these, formatting issues are far too common

\pause
\vspace{.1cm}
\begin{itemize}
\item RDS files (\textit{.rds})
\end{itemize}
$\rightarrow$ Impossible to handle outside of `R`, easy on storage  
$\rightarrow$ I **highly** recommend using these for every step of your work past initial data recording


## Data Handling
### Data Structure
I recommend a structure like the one below with at least two hierarchy levels.  
\vspace{5pt}
\pause
The only files allowed in your first hierarchy level are:
\begin{itemize}
\item R master file
\item Manuscript master file
\end{itemize}

\centering \includegraphics[width=0.8\linewidth]{Figures/Hierarchy.jpg}

\vspace{-.2cm}

\pause
Additionally, make sure to \textbf{back-up your project folder frequently} and use \textbf{version control} on it.

## Data Mining
### The README File
Using the \textbf{README file}, one can identify what information is contained within the data set and thus decide:  
\begin{itemize}
\pause
\item What type/class a data record should be of
\pause
\item Which variables may be redundant
\pause
\item Which data records exceed their variable-specific feasible thresholds
\pause
\item Where to get comparative data sets from
\end{itemize}
\vspace{.5cm}
\pause
\textbf{Data Mining} should then focus on:
\begin{itemize}
\pause
\item \textit{Identifying problems} within the data records
\pause
\item \textit{Explorative} data analyses
\end{itemize}

### Mining in `R` - Numbers or Visualisations?
For data mining, one may wish to enlist the use of Descriptive Statistics \& Data Visualization:
\vspace{10pt}
\begincols[T]
  \begincol{.5\linewidth}
  \pause
  \textbf{Descriptive Statistics:}  
    - `summary()`  
    - `str()`  
    - iterative sub-setting and inspection  
    
  \endcol
  \begincol{.5\linewidth}
  \pause
  \textbf{Data Visualizations:}  
    - Histograms (`hist()`)  
    - Scatter plots (`ggplot2 Package`)  
    
  \endcol
\endcols
\vspace{20pt}
\pause
\centering \textbf{Holistic data mining} is best \textbf{achieved using a combination} of data visualizations tools and parameters of descriptive statistics!

## Data Sharing
### Recording Data Collection - The README File
\textbf{Documenting data recording} is just as important as proper data collection!  
\vspace{10pt}
\pause
To do so, one usually uses a \textbf{README} file containing the following:
\begin{itemize}
  \item Project Name and Summary
  \item Primary contact information
  \item Your name and title (if you aren't the primary contact)
  \item Other people working on the project
  \item Location of data and supporting info
  \item Organization and naming conventions used for the data
  \item Any previous work on the project and where its located
  \item Funding information
\end{itemize}
\pause
\vspace{.5cm}
\centering This file is always \textbf{saved in conjunction with the actual data set}!

### Data Sharing
Open science conduct is essential and you should (read *have to* as a student/employee of Aarhus University) share your data \& coding to ensure **reproducibility** of your work:

\vspace{10pt}
\begincols[T]
  \begincol{.5\linewidth}
  \pause
  \textbf{Peer-to-Peer:}  
    - Raw data  
    - Code  
    - You may just as well point your peers to public repositories  
    
  \endcol
  \begincol{.5\linewidth}
  \pause
  \textbf{Public:}  
    - Raw data  
    - Code  
    - Html visualizations (`shiny`, `mapview`)  
    - Websites
    
  \endcol
\endcols
\vspace{20pt}
\pause
\textbf{Aarhus Guideline:} Store data on the Ecoinf/Biochange data server. NOT on the computational server. Read more \href{https://medarbejdere.au.dk/en/administration/researchandtalent/responsible-conduct-of-research/policy-for-research-integrity-freedom-of-research-and-responsible-conduct-of-research-at-aarhus-university/}{here}.


# Statistical Assumptions
### Assumptions
Statistical tests rely on individual \textit{statistical assumptions}. Most prominent:
\begincols[c]
  \begincol{.5\linewidth}
  \begin{itemize}
  \pause
  \item \textbf{Normality}: Data follow a normal distribution
  \pause
  \item \textbf{Randomness}: Data are truly random
  \pause
  \item \textbf{Independence}: Data are independent
  \pause
  \item \textbf{Homogeneity of variances}: Data from separate groups have same variance
  \pause
  \item \textbf{Linearity}: Data have linear relationship
  \end{itemize}
  \endcol
  \begincol{.5\linewidth}
  \visible<2->{
  \centering \includegraphics[width=1\linewidth]{Figures/FB_IMG_1508796028117.jpg}
  }
  \endcol
\endcols

## Normality
### Normality
\begin{tcolorbox}[colback=burgundy!5,colframe=burgundy!40!black,title=]
\centering Testing for normality of the data is \textbf{crucial} for certain statistical procedures.
\end{tcolorbox}

\begincols[T]
\begincol{.5\linewidth}
  \visible<2->{\textbf{The Shapiro-Wilks Test In Theory}}
  \begin{itemize}
  \visible<3->{
  \item Base assumption: The data is normally distributed
  \item If $p$-value $<$ chosen significance level, the data is \textbf{not} normally distributed
  \item Very sensitive to sample size
  }
  \end{itemize}
\endcol
\begincol{.5\linewidth}
  \visible<2->{\textbf{The QQ Plot In Theory}}
  \begin{itemize}
  \visible<4->{
  \item Method for comparing two probability distributions by plotting their quantiles against each other
  \item If the two distributions being compared are similar, the plot will show the line $y = x$.
  \item Compare the data distribution to the normal distribution
  }
  \end{itemize}
\endcol
\endcols

## Independence
### Independence
\small
\begincols[T]
  \begincol{.5\linewidth}
  \pause
    \textbf{Theory:}
    \begin{itemize}
    \pause
    \item Even the smallest dependence in your data can turn into heavily biased results (which may be undetectable).
    \pause
    \item A dependence is a connection between/within the data.
    \pause
    \item The assumption of independence relies on the absence of any connection in your data that haven't been accounted for in your approach (accounting for it is difficult).
    \end{itemize} 
\endcol
\begincol{.5\linewidth}
  \pause
    \textbf{Independent data:}
    \begin{itemize}
    \pause
    \item \textit{Between Groups}\\
    Groups of data records should be pulled from different individuals.
    \pause
    \item \textit{Within Groups}\\
    Data values within the same group are not to influence one another.
    \pause
    \item  \textit{Within Individuals}\\
    Data values recorded for one individual should not influence each other. This is often an issue with repeated measurement approaches.
    \end{itemize} 
\endcol
\endcols
\vspace{10pt}
\pause
\centering $\rightarrow$ Fixing this \textit{after data collection} is \textbf{almost impossible}!

## Homogeneity of Variances
### Homogeneity of Variances
Particularly important for t-Tests and ANOVAs



\begincols[c]
  \begincol{.5\linewidth}
  \pause
  - **Assumption**: Data from separate groups have same variance  
  - **Test**: `leveneTest()` in the `car` package.  
  
  \endcol
  \begincol{.5\linewidth}
  ```{r homogeneity1, echo = FALSE, fig.height = 6}
  set.seed(42)
  A <- rnorm(n = 1000, 50, 20)
  set.seed(42)
  B <- rnorm(n = 1000, 50, 10)
  boxplot(A, B)
  ```
  \endcol
\endcols

\footnotesize
```{r homogeneity2, echo = FALSE}
library(car)
leveneTest(y = c(A, B), group = as.factor(rep(c("A","B"), each = 1000)))
```