---
title: "Statistical Terminology"
subtitle: "The Basics, Misconceptions, and Pedantises"
author: "Erik Kusch"
date: "04/03/2020"
fontsize: 10pt
output: 
  beamer_presentation: 
    keep_tex: true # keep latex file that is generated during compilation
    toc: false # this is added through a later command
    slide_level: 3 # at how many pound signs (#) to assume slide title level
    includes:
      in_header: Style.tex
classoption: c # change to `handout` to ignore breaks
---
  
```{r setup, include=FALSE}
options(scipen=1, digits=4)
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, cache = FALSE, cache.lazy=TRUE, tidy.opts=list(width.cutoff=50),tidy=TRUE, fig.height=8, digits=4)
```

<!--- ####### TOC (use this if you want ## headers to be included in the toc)--------------------
---------------------------------------  --->
  
\tableofcontents

<!--- ####### DOCUMENT --------------------
---------------------------------------  --->

# Biostatical Terms
## Population vs. Sample
### Population vs. Sample
\begincols[T]
  \begincol{0.5\linewidth}
  \pause
  \textbf{Population:} describes the sum total of all \textit{existing} values of a variable given a certain research question. This includes non-measured data.
  \endcol
  \begincol{0.5\linewidth}
  \pause
  \textbf{Sample:} describes the sum total of all \textit{available} values of a variable for any given analysis. This can only include measured data.
  \endcol
\endcols

\vspace{.2cm}
\pause
\begin{table}[c]
        \centering
          \resizebox{\textwidth}{!}{%
          \begin{tabular}{L{0.2\linewidth}L{0.8\linewidth}}
           \hline
            & \\
          \end{tabular}}
        \label{Spacer}
\end{table}

\vspace{-.7cm}
\textbf{An example:}  \linebreak
In an experimental set-up, you rear an ant colony of exactly 10,000 individuals. You are interested in the average mandible strength of ants within the colony.
 
  \vspace{0.2cm}
  
\pause \textit{The problem:} You cannot possibly take measurements of all 10,000 individuals.

  \vspace{0.2cm}
  
\pause \textit{The solution:} Taking measurements on a \textbf{Sample} (e.g. 1,000 individuals) from within the \textbf{Population} (10,000 individuals).

## Test- vs. Training-Data
### Test- vs. Training-Data
This differentiation is only applicable when concerned with \textit{modelling}.
\vspace{0.2cm}
\begincols[T]
  \begincol{0.5\linewidth}
  \pause
  \textbf{Training Data:} describes the subset of the total data which is used to \textit{establish/train} the model.
  \endcol
  \begincol{0.5\linewidth}
  \pause
  \textbf{Test Data:} describes the subset of the total data which is used to \textit{test} the performance of the model.
  \endcol
\endcols

\vspace{.2cm}
\pause
\begin{table}[c]
        \centering
          \resizebox{\textwidth}{!}{%
          \begin{tabular}{L{0.2\linewidth}L{0.8\linewidth}}
           \hline
            & \\
          \end{tabular}}
        \label{Spacer}
\end{table}

\vspace{-.7cm}
\textit{The problem:} You have identified a way to model how mandible strength and ant size are interconnected but don't know how to assess the quality of your model (a model will always fit the data it was built on extremely well).

\vspace{0.2cm}

\pause \textit{The solution:} Split the available data into two non-overlapping subsets of data (\textbf{Training} and \textbf{Test Data}) and use these separately to build your model and assess its performance.

## Randomness
### Randomness
\begin{tcolorbox}[colback=burgundy!5,colframe=burgundy!40!black,title=]
\centering \textbf{Randomisation} is one of the \textbf{most important} practices in biological studies.
\end{tcolorbox}

\pause

A \textbf{sampling} procedure is \textbf{random} when any member of the \textit{population} has an equal chance of being selected into the \textit{sample}.

\vspace{0.2cm}

\textit{Training} and \textit{Test Data Sets} are established from the population with the same sense of randomness although there may be exceptions depending on the modelling procedure at hand.

\vspace{0.2cm}
\begincols[T]
  \begincol{0.5\linewidth}
  \pause
  \textbf{Data collection:} Number all units contained within the set-up and sample those units corresponding to random numbers.
  \endcol
  \begincol{0.5\linewidth}
  \pause
  \textbf{In R:} Use the `sample()` function to create truly random subsets. Remember to use `set.seed()` to make this step reproducible!
  \endcol
\endcols

### Stratified Sampling
\begin{center}
When do we break \textit{true} \textbf{randomness}?.
\end{center}

\vspace{0.2cm}

\pause

When a \textbf{population} can be divided into distinct categories (i.e. \textbf{strata}). These can be regarded as individual sub-populations.

\vspace{0.2cm}

\textbf{Stratified sampling} ensures that all sub-populations are proportionally represented in the final population-sample given their relative size.

\pause

\scriptsize
\vspace{0.2cm}
\begincols[T]
  \begincol{0.2\linewidth}
```{r stratasetup, echo = FALSE}
s <- c(rep("A",50),rep("B",35),rep("C",15))
d <- as.data.frame(table(s))
```
```{r stratasetup2}
d
```
  \endcol
  \begincol{0.8\linewidth}
```{r strata}
set.seed(42) # stratified
table(sample(d$s, replace = TRUE, prob = d$Freq, 100))
set.seed(42) # non-stratified
table(sample(d$s, replace = TRUE, 100))
```
  \endcol
\endcols


## Supervised vs. Unsupervised Approaches
### Unsupervised Approaches
Unsupervised methods are often \textit{used to select} the \textit{most informative $X$ input variables} for \textit{supervised approaches}. 
\vspace{0.4cm}
\begincols[T]
\begincol{0.5\linewidth}
\pause
\textbf{Pre-requisites:}
\begin{itemize}
\item Only \textit{input variables} are observed.
\item \textit{No solution/feedback (output)} is given. 
\end{itemize}
\endcol
\begincol{0.5\linewidth}
\pause
\textbf{Aims:}
\begin{itemize}
\item \textit{Divide} the observations into relatively distinct groups.
\item \textit{Model} the underlying structure or distribution in the data.
\end{itemize}
\endcol
\endcols
\pause
\vspace{0.8cm}
$\rightarrow$ \textbf{"Pre-processing" before a supervised learning analysis} and \textbf{exploratory analyses}

### Supervised Approaches
Supervised methods are often \textit{informed by unsupervised approaches} and used to \textit{gain validated information} about the data. 
\vspace{0.4cm}
\begincols[T]
\begincol{0.5\linewidth}
\pause
\textbf{Pre-requisites:}
\begin{itemize}
\item Both \textit{predictors} $\mathbf{X}$, and \textit{responses} $\mathbf{Y}$ are observed (there is one $y\textsubscript{i}$ for each $x\textsubscript{i}$).
\item Data is split into \textit{Training} and \textit{Test Data Sets}.
\end{itemize}
\endcol
\begincol{0.5\linewidth}
\pause
\textbf{Aims:}
\begin{itemize}
\item Learn a \textit{mapping function} $f$ from $X$ to $Y$.
\item \textit{Validate} established function/model.
\item Further \textit{prediction} and \textit{inference}.
\end{itemize}
\endcol
\endcols
\pause
\vspace{0.8cm}
$\rightarrow$ \textbf{Mostly inferential analyses}

# Variables \& Scales
## Basics of Variables
### Types of Variables
Variables can be classed into a multitude of types. The most common classification system knows:
\vspace{0.6cm}
\begin{columns}[T]
  \begin{column}{.5\linewidth}
  \pause
    \textbf{Categorical Variables}
    \begin{itemize}
    \item also known as \textit{Qualitative Variables}
    \item Scales can be either:
      \begin{itemize}
      \item Nominal
      \item Ordinal
      \end{itemize}
    \end{itemize}
  \end{column}
  \begin{column}{.5\linewidth}
  \pause
    \textbf{Continuous Variables}
    \begin{itemize}
    \item also known as \textit{Quantitative Variables}
    \item Scales can be either:
      \begin{itemize}
      \item Discrete
      \item Continuous
      \end{itemize}
    \end{itemize}
  \end{column}
\end{columns}

### Categorical Variables
\begin{tcolorbox}[colback=bondiblue!5,colframe=bondiblue!40!black,title=]
\centering Categorical variables are those variables which \textbf{establish and fall into distinct groups and classes}.
\end{tcolorbox}
\pause
\vspace{0.4cm}
Categorical variables:
\begin{itemize}
\item can take on a finite number of values
\item assign each unit of the population to one of a finite number of groups
\item can \textit{sometimes} be ordered
\end{itemize}
\pause
\vspace{0.4cm}
\textbf{In R}, categorical variables usually come up as object type `factor` or `character`.

### Categorical Variables (Examples)
Examples of categorical variables:
\vspace{.5cm}
\begin{itemize}
\pause
\item Biome Classifications (e.g. "Boreal Forest", "Tundra", etc.)
\pause
\item Sex (e.g. "Male", "Female")
\pause
\item Hierarchy Position (e.g. "$\alpha$-Individual", "$\beta$-Individual", etc.)
\pause
\item Soil Type (e.g. "Sandy", "Mud", "Permafrost", etc.)
\pause
\item Leaf Type (e.g. "Compound", "Single Blade", etc.)
\pause
\item Sexual Reproductive Stage (e.g. "Juvenile", "Mature", etc.)
\pause
\item Species Membership
\pause
\item Family Group Membership
\pause
\item ...
\end{itemize}

### Continuous Variables
\begin{tcolorbox}[colback=bondiblue!5,colframe=bondiblue!40!black,title=]
\centering Continuous variables are those variables which \textbf{establish a range of possible data values}.
\end{tcolorbox}
\pause
\vspace{0.4cm}
Continuous variables:
\begin{itemize}
\item can take on an infinite number of values
\item can take on a new value for each unit in the set-up
\item can \textit{always} be ordered
\end{itemize}
\pause
\vspace{0.4cm}
\textbf{In R}, continuous variables usually come up as object type `numeric`.

### Continuous Variables (Examples)
Examples of categorical variables:
\vspace{.5cm}
\begin{itemize}
\pause
\item Temperature
\pause
\item Precipitation
\pause
\item Weight
\pause
\item pH
\pause
\item Altitude
\pause
\item Group Size
\pause
\item Vegetation Indices
\pause
\item Time
\pause
\item ...
\end{itemize}

### Converting Variable Types
\textit{Continuous variables} can be converted into \textit{categorical variables} via a method called \textbf{binning:}

\vspace{0.1cm}

\pause
Given a variable range, one can establish however many "bins" as one wants. For example:
\pause
\begin{itemize}
\item Given a temperature range of $271K - 291K$, there may be 4 bins of equal size:
  \begin{itemize}
  \item Bin A: $271K \leq X \leq 276K$
  \item Bin B: $276K < X \leq 281K$
  \item Bin C: $281K < X \leq 286K$
  \item Bin D: $286K < X \leq 291K$
  \end{itemize}

\end{itemize}
\vspace{0.2cm}
\pause
\begin{tcolorbox}[colback=burgundy!5,colframe=burgundy!40!black,title=]
\centering Whilst a \textbf{continuous variable} can be both \textit{continuous} and \textit{categorical}, a \textbf{categorical variable} can only ever be \textit{categorical}!
\end{tcolorbox}

## Variables And Scales
### Variables On Scales
Another way of classifying variables are the \textbf{scales} they are represented on.
\vspace{0.4cm}
\pause
\begin{tcolorbox}[colback=burgundy!5,colframe=burgundy!40!black,title=]
\centering \textbf{Different scales} of variables \textbf{require different statistical procedures} for analyses!
\end{tcolorbox}
\pause 
\vspace{0.4cm}
Variable scales include:
\begincols[T]
  \begincol{0.5\linewidth}
  \begin{itemize}
  \item \textbf{Nominal} 
  \item \textbf{Binary}
  \item \textbf{Ordinal}
  \end{itemize}
  \endcol
  \begincol{0.5\linewidth}
  \begin{itemize}
  \item \textbf{Interval}
  \item \textbf{Relation/Ratio}
  \end{itemize}
  \endcol
\endcols
\pause 
\vspace{0.4cm}
Some statistics books teach \textit{integer scales} along the above mentioned scales. Some people dispute this and claim these scales to be \textit{ratio scales}.

### Nominal And Binary
\textbf{Nominal scales} of variables correspond to \textit{categorical variables} which cannot be put into a meaningful order.
\begin{columns}[T]
  \begin{column}{.7\linewidth}
    \vspace{0.2cm}
    \pause
    \begin{itemize}
    \item Variables on nominal scales put units into distinct categories
    \item These variables may be numerical but offer no mathematical interpretation
    \end{itemize}
    \pause
    \textit{Examples:}
    \begin{itemize}
    \item Petal colour (red, green, blue, etc.)
    \item Individual IDs
    \end{itemize}
  \end{column}
  \begin{column}{.3\linewidth}
    \begin{center}
    \includegraphics[width=0.65\linewidth]{Figures/Nominal.png}  
    \end{center}
  \end{column}
\end{columns}
\vspace{0.2cm}
\pause
\textbf{Binary scales} are a special case of \textit{nominal scales} taking only two possible values: 0 and 1.

### Ordinal
\textbf{Ordinal scales} of variables correspond to \textit{categorical variables} which can be put into meaningful order.
\begin{columns}[T]
  \begin{column}{.7\linewidth}
    \vspace{0.2cm}
    \pause
    \begin{itemize}
    \item Variables on ordinal scales put units into distinct categories
    \item These variables may be numerical and offer some mathematical interpretation
    \end{itemize}
    \pause
    \textit{Examples:}
    \begin{itemize}
    \item Size (small, medium, large, etc.)
    \item Binned continuous variables
    \end{itemize}
  \end{column}
  \begin{column}{.3\linewidth}
    \vspace{0.2cm}
    \begin{center}
    \includegraphics[width=0.65\linewidth]{Figures/Ordinal.png}  
    \end{center}
  \end{column}
\end{columns}

### Interval/Discrete
\textbf{Interval scales} of variables correspond to a mix of \textit{continuous variables}.
\begin{columns}[T]
  \begin{column}{.7\linewidth}
    \vspace{0.2cm}
    \pause
    \begin{itemize}
    \item Variables on interval scales are measured on equal intervals from a defined zero point/point of origin
    \item The point of origin \textbf{does not imply an absence of the measured characteristic}
    \end{itemize}
    \pause
    \textit{Examples:}
    \begin{itemize}
    \item Temperature [\textdegree$C$]
    \item pH
    \end{itemize}
  \end{column}
  \begin{column}{.3\linewidth}
    \vspace{0.2cm}
    \begin{center}
    \includegraphics[width=0.65\linewidth]{Figures/Interval.png}  
    \end{center}
  \end{column}
\end{columns}

### Relation/Ratio
\textbf{Relation/Ratio scales} of variables correspond to \textit{continuous variables}.
\begin{columns}[T]
  \begin{column}{.7\linewidth}
     \vspace{0.2cm}
    \pause
    \begin{itemize}
    \item Variables on relation/ratio scales are measured on equal intervals from a defined zero point/point of origin
    \item The point of origin \textbf{does imply an absence of the measured characteristic}
    \end{itemize}
    \pause
    \textit{Examples:}
    \begin{itemize}
    \item Temperature [$K$]
    \item Weight
    \end{itemize}
  \end{column}
  \begin{column}{.3\linewidth}
    \begin{center}
    \includegraphics[width=0.65\linewidth]{Figures/Relation.png}  
    \end{center}
  \end{column}
\end{columns}
\pause
\vspace{0.5cm}
\textbf{Integer scales} are a special case of \textit{ratio scales} allowing only for integral numbers.

### Confusion Of Units
\begin{center}
\includegraphics[width=0.5\linewidth]{Figures/Metrics.jpg}  
\end{center}

# Distributions
## The Basics of Distributions
### What Are Distributions?
\begincols[c]
  \begincol{0.5\linewidth}
  \begin{tcolorbox}[colback=bondiblue!5,colframe=bondiblue!40!black,title=]
  \centering A distribution of a statistical data set (sample/population) shows all the possible values/intervals of the data in question and their frequency.
  \end{tcolorbox}
  \endcol
  \begincol{0.5\linewidth}
  \visible<2->{
  \begin{center}
  \includegraphics[width=0.9\linewidth]{Figures/Distris.png} 
  \end{center}
  }
  \endcol
\endcols
\pause
\vspace{1cm}
$\rightarrow$ Basically, \textbf{data patterns} we are considering/looking for.


### Frequency Distributions
\begincols[T]
  \begincol{0.5\linewidth}
  \textbf{Frequency} Distributions:
  \begin{itemize}
  \item \textit{Theory}
    \begin{itemize}
    \item Simple representations of data value frequencies
    \item Can be established for every variable
    \end{itemize}
    \pause
  \item \textit{Practice in R}
    \begin{itemize}
    \item Visualisation via the `hist()` function
    \end{itemize}
  \end{itemize}
\endcol
\begincol{0.5\linewidth}
```{r freqdist_plot, echo=FALSE}
hist(rnorm(100000,20,2),
main = "Frequency Distribution")
```
\endcol
\endcols
```{r freqdist, eval=FALSE, tidy.opts=list(width.cutoff=55), tidy=FALSE}
hist(rnorm(100000,20,2),
main = "Frequency Distribution")
```

### Probability Density Distributions I
\begincols[T]
  \begincol{0.5\linewidth}
  \textbf{Probability Density} Distributions:
  \begin{itemize}
  \item \textit{Theory}
    \begin{itemize}
    \item Representation of data value probabilities
    \item Can be established for \textit{continuous} variables
    \end{itemize}
    \pause
  \item \textit{Practice in R}
    \begin{itemize}
    \item Visualisation via the `density()` function
    \end{itemize}
  \end{itemize}
\endcol
\begincol{0.5\linewidth}
```{r densitydist_plot, echo=FALSE}
plot(density(rnorm(100000,20,2)),
main = "Probability Density Distribution")
```
\endcol
\endcols
```{r densitydist, eval=FALSE, tidy.opts=list(width.cutoff=55), tidy=FALSE}
plot(density(rnorm(100000,20,2)),
  main = "Probability Density Distribution")
```

### Probability Density Distributions II
\begin{tcolorbox}[colback=bondiblue!5,colframe=bondiblue!40!black,title=]
\centering Probability Density Distributions hold the \textbf{majority of importance} in statistics!
\end{tcolorbox}
\vspace{0.4cm}
\pause
A few key points about these distributions:
\begin{itemize}
\pause
\item Area under the curve (AUC) sums to 1
\pause
\item A probability for every given single value is 0
\pause
\item The AUC between two values on the X-axis equals the probability to randomly sample a value between these two points
\end{itemize}

\vspace{.3cm}
\pause
Find a masterful explanation of the single-value probability \href{https://www.youtube.com/watch?v=ZA4JkHKZM50}{here}.

## Normality
### Univariate Standard Normal/Gaussian Distribution
\begin{tcolorbox}[colback=bondiblue!5,colframe=bondiblue!40!black,title=]
\centering One of the \textbf{most important} distributions in natural sciences. 
\end{tcolorbox}

\begincols
  \begincol{.5\linewidth}
  \vspace{-0.2cm}
  \begin{itemize}
  \pause
  \item Used to represent real-valued random variables whose distributions are not known
  \pause
  \item The \textbf{central limit theorem} applies (draw a sufficient number of samples and you end up with the normal distribution)
  \pause
  \item These distributions are usually known also as "bell curves" (\textbf{Attention:} other distributions take this shape too)
  \end{itemize}
  \endcol
  \begincol{.5\linewidth}
  \visible<4->{
  \begin{center}
  \includegraphics[width=0.6\linewidth]{Figures/Normal.jpg}  
  \end{center}
  }
  \endcol
\endcols

### Testing For Normality
\begin{tcolorbox}[colback=burgundy!5,colframe=burgundy!40!black,title=]
\centering Testing for normality of the data is \textbf{crucial} for certain statistical procedures.
\end{tcolorbox}

\begincols[T]
\begincol{.5\linewidth}
  \visible<2->{\textbf{The Shapiro-Wilks Test In Theory}}
  \begin{itemize}
  \visible<3->{
  \item Base assumption: The data is normally distributed
  \item If $p$-value $<$ chosen significance level, the data is \textbf{not} normally distributed
  \item Very sensitive to sample size
  }
  \end{itemize}
\endcol
\begincol{.5\linewidth}
  \visible<2->{\textbf{The QQ Plot In Theory}}
  \begin{itemize}
  \visible<4->{
  \item Method for comparing two probability distributions by plotting their quantiles against each other
  \item If the two distributions being compared are similar, the plot will show the line $y = x$.
  \item Compare the data distribution to the normal distribution
  }
  \end{itemize}
\endcol
\endcols


### The Shapiro-Wilks Test In R
Using the `shapiro.test()` function:

\footnotesize
\begincols[T]
\pause
\begincol{.5\linewidth}
```{r shapiroyes}
shapiro.test(rnorm(5000,20,2))
```
$\rightarrow$ Clearly a normal distributed set of values
\endcol
\begincol{.5\linewidth}
\pause
```{r shapirono}
shapiro.test(seq(1,500,5))
```
$\rightarrow$ Clearly no normal distributed set of values
\endcol
\endcols

\normalsize
\pause
\vspace{.4cm}
For data sets bigger than 5000 data points, use the Kolmogorov-Smirnov test (`ks.test()`) in `R`.


### The Q-Q Plot
Using the `qqnorm()` function:
\footnotesize
\begincols
\begincol{.5\linewidth}
\pause
```{r QQyes, tidy=FALSE}
qqnorm(rnorm(5000,20,2))
```
$\rightarrow$ Clearly a normal distributed set of values
\endcol
\begincol{.5\linewidth}
\pause
```{r QQno, tidy=FALSE}
qqnorm(seq(1,500,5))
```
$\rightarrow$ Clearly no normal distributed set of values
\endcol
\endcols


## What Distributions To Consider
### Binomial Distribution
One of the **more important** distributions. It is applicable to:  

\begincols[T]
  \begincol{.5\linewidth}
  \pause
  
  * Variables which can only take two possible values (e.g. "states")  
  * All records of the variable have the same probability *p* of being in one of the two states  
  
  \pause
  It is made up of three **criteria**:  
  \pause
  
  \begin{itemize}
  \item $p$ - the "success" probability  
  \item $n$ - sample size (how often we sample)  
  \item $N$ - the "binomial total" (for how many individuals we sample each time)
  \end{itemize}
  
  \endcol
  \begincol{.5\linewidth}
  ```{r distbinom, echo = FALSE}
  set.seed(42)
  hist(rbinom(n = 10000, size = 50, prob = 0.6), main = "Binomial Distribution", xlab = 
  "n = 10000, N = 50, p = 0.6")
  ```
  \endcol
\endcols

### Poisson Distribution
Another one of the **more important** distributions. It is applicable to:  

\begincols[T]
  \begincol{.5\linewidth}
  \pause
  
  * Focal objects are placed randomly in one or more dimensions  
  * A random "counting window" (usually one considering time) is placed above the sampling scheme  
  
  \pause
  It is made up of two **criteria**:  
  \pause
  
  \begin{itemize}
  \item $\lambda$ - the mean (= expectation, average count, intensity) as well as
the variance (i.e., variance = mean)
  \item $n$ - sample size
  \end{itemize}

  \endcol
  \begincol{.5\linewidth}
  ```{r distpoisson, echo = FALSE}
  set.seed(42)
  hist(rpois(1000, 5), main = "Poisson Distribution", xlab = 
  "n = 10000, Lambda = 5")
  ```
  \endcol
\endcols

## Important Measures Of Distributions
### How to Measure Distributions
\begin{center}
\textit{Not all distributions are created equally.}
\end{center}
\vspace{.3cm}
\pause
Distributions can be described via \textbf{classic parameters of descriptive statistics:}
\begin{columns}[T]
  \begin{column}{.5\linewidth}
    \begin{itemize}
    \pause
    \item Arithmetic Mean
    \item Mode
    \item Median
    \item Minimum, Maximum, Range
    \item ...
    \end{itemize}
  \end{column}
  \begin{column}{.5\linewidth}
    \begin{itemize}
    \pause
    \item Variance
    \item Standard Deviation
    \item Quantile Range
    \item \textbf{Skewness}
    \item \textbf{Kurtosis}
    \item ...
    \end{itemize}
  \end{column}
\end{columns}

### Skewness I

\begin{table}[c]
        \centering
          \resizebox{\textwidth}{!}{%
          \begin{tabular}{L{0.2\linewidth}L{0.8\linewidth}}
           \textit{Definition:} & Describes the symmetry and relative tail length of distributions. \\
             & \\
           \hline
             & \\
           \textit{Positive skew:} & Right-hand tail is longer than the left-hand tail\\
           \textit{Skew = 0:} & Symmetric distribution\\
           \textit{Negative skew:} & Left-hand tail is longer than the right-hand tail\\
          \end{tabular}}
        \label{Des_Stats}
\end{table}


### Skewness II
\vspace{.6cm}
\begin{columns}[T]
  \begin{column}{0.3\linewidth}
    \begin{center}
    \includegraphics[width=1\linewidth]{Figures/PosSkew.png}  
    \end{center}
    \centering\tiny{\textit{Positive Skew}}
  \end{column}
  \begin{column}{.4\linewidth}
    \begin{center}
    \includegraphics[width=1\linewidth]{Figures/Symm.png}  
    \end{center}
    \centering\tiny{\textit{Symmetric Distribution}}
  \end{column}
  \begin{column}{0.3\linewidth}
    \begin{center}
    \includegraphics[width=1\linewidth]{Figures/NegSkew.png}  
    \end{center}
    \centering\tiny{\textit{Negative Skew}}
  \end{column}
\end{columns}


### Kurtosis I
\begin{table}[c]
        \centering
          \resizebox{\textwidth}{!}{%
          \begin{tabular}{L{0.2\linewidth}L{0.8\linewidth}}
           \textit{Definition:} & Describes the evenness/"tailedness" of distributions. \\
             & \\
           \hline
             & \\
           \textit{Positive kurtosis:} & Short-tailed distribution aka. \textit{leptokurtic}\\
           \textit{Kurtosis = 0:} & Base representation of a given distribution aka. \textit{mesokurtic}\\
           \textit{Negative kurtosis:} & Long-tailed distribution aka. \textit{platykurtic}\\
          \end{tabular}}
        \label{Des_Stats}
\end{table}

### Kurtosis II
\vspace{.6cm}
\begin{center}
\includegraphics[width=0.8\linewidth]{Figures/Kurtosis.jpg}  
\end{center}