---
title: "Troubleshooting R"
subtitle: "Isolating Issues and Asking Questions"
author: "Erik Kusch"
date: "17/02/2021"
fontsize: 10pt
output: 
  beamer_presentation: 
    keep_tex: true # keep latex file that is generated during compilation
    toc: false # this is added through a later command
    slide_level: 3 # at how many pound signs (#) to assume slide title level
    includes:
      in_header: Style.tex
classoption: c # change to `handout` to ignore breaks
---
  
```{r setup, include=FALSE}
options(scipen=1, digits=4)
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, cache = TRUE, cache.lazy=TRUE, tidy.opts=list(width.cutoff=50),tidy=FALSE, fig.height=8, digits=4)
```

<!--- ####### TOC (use this if you want ## headers to be included in the toc)--------------------
---------------------------------------  --->
  
\tableofcontents

<!--- ####### DOCUMENT --------------------
---------------------------------------  --->

# Dealing with Problems
## Problem-Sources
### So your analysis is misbehaving...
\vspace{-.2cm}
\begin{tcolorbox}[colback=burgundy!5,colframe=burgundy!40!black,title=]
\centering Start by \textbf{investigating common errror sources}:
\end{tcolorbox}
\vspace{-.2cm}

\pause
\begincols[T]
  \begincol{.33\linewidth}
    \begin{center}
    \textbf{Code Errors}
    \end{center}
    \visible<3->{
    \begin{itemize}
    \item Capitalisation
    \item Mismatched... 
      \begin{itemize}
      \item Object types
      \item Value modes
      \item Object dimensions
      \end{itemize}
    \item Directory structures
    \item Non-self-contained scripts
    \end{itemize}
    }
  \endcol
  \begincol{.33\linewidth}
    \begin{center}
    \textbf{Mis-Specified Statistics}
    \end{center}
    \visible<4->{
    \begin{itemize}
    \item Assumptions not met
    \item Erroneous outliers
    \item Mis-interpretation of outputs
    \item Incorrect data handling
    \end{itemize}
    }
  \endcol
  \begincol{.33\linewidth}
    \begin{center}
    \textbf{Unexpected Findings}
    \visible<5->{
    \textit{"[...] in an experimental framework, surprising results are significant results, rather than signs of failure."} \linebreak
    $\sim$ Curtin \& Parker, 2014
    }
    \end{center}
  \endcol
\endcols
\visible<6>{
\begin{tcolorbox}[colback=green!5,colframe=green!40!black,title=]
\centering How do we resolve these?
\end{tcolorbox}
}


## Approaching Problems
### Errors and Warnings
\begincols[T]
  \begincol{.5\linewidth}
    `Error` \linebreak
    `R`: "I can't and won't. You won't force me. I quit. I am done." \linebreak
    \centering \includegraphics[width=.9\linewidth]{Figures/error.png}
  \endcol
  \begincol{.5\linewidth}
    `Warning` \linebreak
    `R`: "I can, but you might not like the output better look at that, my guy." \linebreak
    \centering \includegraphics[width=.9\linewidth]{Figures/warning.png}
  \endcol
\endcols

\vspace{.2cm}
\raggedleft \tiny \url{http://rex-analytics.com/decoding-error-messages-r/}

### Breaking problems down
<!-- \textit{"If you're overwhelmed by the size of a problem, break it down into smaller pieces."} \linebreak -->
<!-- \raggedleft $\sim$ Chuck Close -->
- Investigate objects and their contents  
- Read function help files (especially the arguments and values sections)  
- Simulate data if in doubt  
\hfill \linebreak
\visible<2->{
\centering \includegraphics[width=.75\linewidth]{Figures/eggs.jpg}
\raggedleft \tiny Sousa et al., 2007 \normalsize
}
\visible<3->{
\begin{tcolorbox}[colback=green!5,colframe=green!40!black,title=]
\centering Your analyses hardly ever fail at all steps.
\end{tcolorbox}
}



# Reproducibiliy
## What & Why
### What it is and why you need to care

**Reproducibility** ensures:  

- Honest and true reporting  
- Results can be scrutinised  
- Problems can be traced and identified  
- Co-authors and reviewers get less annoyed  

\vspace{.4cm}
\begin{tcolorbox}[colback=green!5,colframe=green!40!black,title=]
\textbf{Our Goal:} \linebreak
\centering An analysis is \textbf{reproducible} when it can be executed by different researchers across operating systems with a \textbf{single button prompt} and return the same results for everyone.
\end{tcolorbox}

## How
### Coding Styles
- **Code**  
    - *Consistency*  
        - The same result on every execution.  
        - The same naming scheme throughout.  
    - *Self-contained* - Code runs in an empty `R` environment to completion.  

\vspace{.4cm}    

- **Documentation**  
    - *Commenting* - Informative comments for:  
        - What each line is doing/supposed to do.  
        - Why each line is doing what it does/is supposed to do.  
    - *Header* - Information of what the script does, and who authored it.  
    - *File-Versioning* - GitHub or an equivalent to trace when a script broke.  

### Packages

\begin{tcolorbox}[colback=burgundy!5,colframe=burgundy!40!black,title=]
\centering \textbf{Never include} forced \textbf{package installation} in a script.
\end{tcolorbox}
\vspace{.3cm}

Instead, do something like this:
\footnotesize
```{r, eval = FALSE}
install.load.package <- function(x) {
  if (!require(x, character.only = TRUE))
    install.packages(x, repos='http://cran.us.r-project.org')
  require(x, character.only = TRUE)
}
package_vec <- c(
  "rethinking", # for quadratic approximation of posteriors
  "reshape2" # for data reformatting
)
sapply(package_vec, install.load.package)
```

### Working Directories
\begin{tcolorbox}[colback=burgundy!5,colframe=burgundy!40!black,title=]
\centering \textbf{Never} include \textbf{hard-coded} directory indexing.
\end{tcolorbox}
\only<1-6>{
\begincols[T]
  \begincol{.5\linewidth}
    \only<1-6>{
    \textbf{Make an Omlette out of Easter Eggs} \linebreak
    \begin{enumerate}
    \item Find the egg(s)/data
    \item Open the egg(s)/data
    \item Make the egg(s) into an omelette / do your analysis
    \end{enumerate}
    \pause
    .... but... where are the eggs/the data?
    }
  \endcol
  \begincol{.5\linewidth}
    \only<3>{\centering \includegraphics[width=.9\linewidth]{Figures/C.png}}
    \only<4>{\centering \includegraphics[width=.9\linewidth]{Figures/User.png}}
    \only<5-6>{\centering \includegraphics[width=.9\linewidth]{Figures/Project.png}}
  \endcol
\endcols

\vspace{.4cm}
\visible<6>{
\centering But what if you want to hunt for Easter Eggs/data somewhere other than the garden at Achmore/your specific project folder?
}
}

\only<1-6>{\vspace{1cm}}
\only<7>{\vspace{.1cm}}
Soft-code your working directories from a base/project directory:
\footnotesize
```{r}
Dir.Base <- getwd() # to find the project folder
Dir.Data <- file.path(Dir.Base, "Data") # index the data folder
```

\normalsize

Now we have our directories indexed without any hardcoding:

\tiny
```{r}
Dir.Base
Dir.Data
```

### Randomness
\begin{tcolorbox}[colback=burgundy!5,colframe=burgundy!40!black,title=]
\centering \textbf{Always} make \textbf{random processes reproducible}.
\end{tcolorbox}
\only<1-4>{
\begincols[T]
  \begincol{.65\linewidth}
  \only<1-4>{
  Imagine you survey the elevation of a random surface with paratroopers:
    \begin{itemize}
    \item Clouds obscure the landscape
    \item All flights are on the same route
    \item Troopers report elevation of their landing
    \item There is no drift of paratroopers
    \item Troopers jump at fixed intervals starting at a certain time after take-off
    \end{itemize}
  }
  \endcol
  \begincol{.35\linewidth}
    \only<1>{\centering \includegraphics[width=1\linewidth]{Figures/Para1.jpg}}
    \only<2-4>{\centering \includegraphics[width=1\linewidth]{Figures/Para2.jpg}}
    \only<4>{
$\rightarrow$ We let all flights commence their jump at the same time after take-off.
}
  \endcol
\endcols

\normalsize
\vspace{.4cm}
\visible<3-4>{
$\rightarrow$ How can we get the same elevations from the troopers of two different flights?
}
}
\only<1-4>{\vspace{1cm}}
\only<5>{\vspace{.1cm}}
\footnotesize
```{r}
rnorm(5, 1, 0.2)
rnorm(5, 1, 0.2)
set.seed(42); rnorm(5, 1, 0.2)
set.seed(42); rnorm(5, 1, 0.2)
```

\begin{tcolorbox}[colback=green!5,colframe=green!40!black,title=]
\centering A seed needs to be set before \textbf{every} random process.
\end{tcolorbox}


# Reporting Issues and Asking for Help
## Minimal Working Examples
### What is a Minimal Working Example (MWE)?  

\begin{tcolorbox}[colback=green!5,colframe=green!40!black,title=]
\centering A reproducible, simplified version of code producing a desired outcome.
\end{tcolorbox}

- **Minimal**. Information and code is reduced as much as possible. This involves reducing:   
    - Lines of `R` code  
    - Number of `R` packages needed  
    - Reliance on directory structures  
    - Description of work, result, aim, and potential error  
- **Complete**. All information, data, and code necessary to execute the MWE is presented.    
- **Reproducible**. The MWE returns the same output on all operating systems when executed by any person. Ideally, this happens at the push of a single button.  


## Minimality
### Minimal Description

**The Goal:** *I want to use a k-means clustering approach to identify biome classes across Alaska during the year 1982 using GIMMs NDVI 3g annual mean and seasonality values.*

\vspace{.3cm}

**Method(s) & Material(s):**  
- *Data:*  
    - GIMMs NDVI 3g (9x9km index of vegetation performance)  
    - Natural Vector Shapefiles (shapefile-collection of states and provinces)  
- *Method:*  
    - K-Means-Clustering implemented via the `mclust` package

\vspace{.3cm}
    
**The Problem:**  Call to `mclustBIC()` produces the following error:  

`Error in mvnXII(data = data, prior = prior) : NA/NaN/Inf in foreign function call (arg 1)`

### Minimal Data - Before Reducing to MWE

**The Data on our end** before handing it off:  
\begincols[T]
  \begincol{.5\linewidth}
  \centering \includegraphics[width=1\linewidth]{Figures/DataFull.JPG}
  \endcol
  \begincol{.5\linewidth}
  \centering \includegraphics[width=1\linewidth]{Figures/DataFullHDD.JPG}
  \endcol
\endcols

\vspace{.3cm}

Data needs to be **reduced** to the important parts (`All1982_ras`, in this case). We can export an `R` environment with the `save.image()` function.

\vspace{.3cm}

**The Data for the MWE** that we hand over:  
\begincols[T]
  \begincol{.5\linewidth}
  \centering \includegraphics[width=1\linewidth]{Figures/DataSparse.JPG}
  \endcol
  \begincol{.5\linewidth}
  \centering \includegraphics[width=1\linewidth]{Figures/DataSparseHDD.JPG}
  \endcol
\endcols



### Minimal Code

\begincols[T]
  \begincol{.5\linewidth}
  \textbf{Original Script:}
  \centering \includegraphics[width=1\linewidth]{Figures/ErrorScript.JPG}
  \endcol
  \begincol{.5\linewidth}
  \textbf{MWE Script:}
  \centering \includegraphics[width=1\linewidth]{Figures/MWEScript.JPG}
  \vspace{.3cm}
  We removed code for:
  \begin{itemize}
  \item Download of data
  \item Cropping and masking of data
  \item Calculation of compound metrics
  \end{itemize}
  \endcol
\endcols

## Completeness
### Complete Reporting

\begincols[T]
  \begincol{.5\linewidth}
  We also supply:
  \begin{itemize}
  \item Archive of our raw data directory
  \item Script which contains all the code we removed from full script when creating MWE script
  \end{itemize}
  \vspace{.3cm}
We may also want to provide:  
- `sessionInfo()`  
- `rstudioapi::versionInfo()`  
- Metadata if needed  
  \endcol
  \begincol{.5\linewidth}
  \textbf{MWE Script:}
  \centering \includegraphics[width=1\linewidth]{Figures/SupplementScript.JPG}
  \endcol
\endcols

## Reproducibility
### Reproducible Issues  

\begincols[T]
  \begincol{.5\linewidth}
We test our MWE script:  
- In a new, empty directory by sourcing the script  
- For completeness by running it on an empty `R` environment and by adding `rm(list=ls())` to the start of the script  

\vspace{.3cm}

Our script still produces the same error?  

$\rightarrow$ We ship it off to someone whom we hope can help us.
  \endcol
  \begincol{.5\linewidth}
  \centering \includegraphics[width=1\linewidth]{Figures/908.jpg}
  \tiny \raggedleft \url{https://knowyourmeme.com/photos/1297214-impossible-perhaps-the-archives-are-incomplete}
  \endcol
\endcols





