---
title: "Rmarkdown - Manuscript Workflow Revisited"
subtitle: "Using RStudio to Write Theses, Paper, and Presentations"
author: "Erik Kusch"
date: "10/09/2020"
fontsize: 10pt
output: 
  beamer_presentation: 
    keep_tex: true # keep latex file that is generated during compilation
    toc: false # this is added through a later command
    slide_level: 3 # at how many pound signs (#) to assume slide title level
    includes:
      in_header: Style.tex
classoption: t # change to `handout` to ignore breaks
---
  
```{r setup, include=FALSE}
options(scipen=1, digits=4)
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, cache = TRUE, cache.lazy=TRUE, tidy.opts=list(width.cutoff=50),tidy=TRUE, fig.height=8, digits=4)
```

<!--- ####### TOC (use this if you want ## headers to be included in the toc)--------------------
---------------------------------------  --->
  
\tableofcontents[subsubsectionstyle=hide]

<!--- ####### DOCUMENT --------------------
---------------------------------------  --->

# Introduction
### What is `Rmarkdown`?
\begincols
  \begincol{.5\linewidth}
  \vspace{-.53cm}
  \pause
  **What it is** (*`Rmarkdown`*/*`.Rmd`*) \newline
  \pause
  In its **essence**, `Rmarkdown` is a **coding document containing text** that can be used to generate:
  \begin{itemize}
  \item Manuscripts
  \item Presentations
  \item Posters
  \item Books
  \item ...
  \end{itemize}
  whilst **maintaining reproducibility** of research. \newline
  \pause
  It is *heavily reliant* on the \LaTeX \hspace{.04cm} machinery.
  \endcol
  \begincol{.5\linewidth}
  \pause
  **What you need** (*Dependencies*) \newline
  To be able to use the bare minimum of `Rmarkdown`, you need an installation `R` and `RStudio` (These are the *basics*). \newline
  \pause
  *Advanced* use of `Rmarkdown` is usually funelled through \LaTeX \hspace{.04cm} and so you'll need installations of:  
  \begin{itemize}
  \item \LaTeX \hspace{.04cm} (for generation of pdf documents)
  \item JabRef (for referencing)
  \item Perl (for glossaries)
  \end{itemize}
  \endcol
\endcols

### Why use `Rmarkdown`?
Using `Rmarkdown` for your research comes with a multitude of advantages:  \vspace{.3cm}

1. Entire **workflow in one program** (`RStudio`)  
\vspace{.1cm}
2. **Research** and reports **reproducible** at the click of **one button**  
\vspace{.1cm}
3. **Combines** `R` functionality and \LaTeX \hspace{.04cm} formatting (if desired)  
\vspace{.1cm}
4. **Consistent formatting**  
\vspace{.1cm}
5. **Clear presentation of code**  
\vspace{.1cm}
6. **Dynamic documents** (you can generate various output document types)  
\vspace{.1cm}
7. Applicable for **almost all document types** you may desire as an output (e.g. manuscripts, presentations, posters, etc.)  

# The Basics
## Getting Started
### `Rmarkdown` Worflow
In the file menu of `RStudio`, click: **File** $\rightarrow$ **New File** $\rightarrow$ **R Markdown**
\vspace{.1cm}
\begin{center}
  \includegraphics[width=0.6\linewidth]{Figures/Open.jpg}
\end{center}
\vspace{.1cm}
Write and edit the document and *compile* by **knitting**.

### `Rmarkdown` Resources
I cannot cover every aspect of `Rmarkdown` coding in one go. In case you bite, here are resources to help you along:
\vspace{.2cm}
\begin{itemize}
\pause
  \item \textbf{Guides}
    \begin{itemize}
    \item An Introduction: \url{https://rmarkdown.rstudio.com/lesson-1.html}
    \item The \textit{'Definitive Guide'}: \url{https://bookdown.org/yihui/rmarkdown/}
    \end{itemize}
    \vspace{.2cm}
  \item \textbf{CheatSheets}
  \pause
  \begin{itemize}
    \item Reference Guide: \url{https://www.rstudio.com/wp-content/uploads/2015/03/rmarkdown-reference.pdf}
    \item Proper CheatSheet: \url{https://www.rstudio.com/wp-content/uploads/2016/03/rmarkdown-cheatsheet-2.0.pdf}
    \end{itemize}
\end{itemize}
\vspace{.2cm}
\pause
$\rightarrow$ I recommend learning `Rmarkdown` by inspecting working scripts (including the one used to produce this presentation). You can find some on my github: \url{https://github.com/ErikKusch}.

## `Rmarkdown` Document Structure
### Overview
`Rmarkdown` files usually come down to \textbf{three main components}:
\vspace{.3cm}
\begin{enumerate}
\pause
\item The \textit{YAML Header} - This sets up important characteristics of your document such as:
  \begin{itemize}
  \item Title, Subtitle
  \item Author and Date
  \item Output File Type
  \item Document Class and Packages (if output type is \LaTeX \hspace{.04cm} dependant)
  \end{itemize}
\vspace{.15cm}
\pause
\item \textit{Plain Text} - This is where you write the text of your document. \LaTeX \hspace{.04cm} formatting is supported when the YAML Header indicates a \LaTeX \hspace{.04cm} dependant output file type.
\vspace{.15cm}
\pause
\item \textit{Code Chunks} - Here, you will implement R codes (Matlab and Python coding is supported as well) and run your analyses.
\end{enumerate}

### YAML Header

A non-exhaustive list of YAML specifiers:

YAML | Function
------------- | -------------
title | The heading of your work
subtitle | A subheading for your work
author | You and co-workers (if applicable)
date | Date of delivery or presentation
output | Whether to create html, pdf, or presentation
includes | Packages and auxiliary files
documentclass | \LaTeX \hspace{.04cm} document class (if applicable)
classoption | \LaTeX \hspace{.04cm} document option (if applicable)
fontsize | Base font size throughout document


### Text

**Formatting** can offset the inherently boring nature of plain text and can be done:
\vspace{.1cm}
\begin{itemize}
\item Through base syntax
\item Through \LaTeX \hspace{.04cm} syntax
\end{itemize}

\pause
\vspace{.5cm}
Text is *hard coded* but results and code outputs can be included through *soft-coding* via **in-line code**.

\pause
\vspace{.5cm}
Text can also be supplied via **auxillary files** using the `child` syntax supplied by `Rmarkdown`. This is useful for big documents but makes cross-referencing very complicated.

### Code Chunks

`R` Code chunks are started with \textasciigrave\textasciigrave\textasciigrave\textit{\{r\}} and ended with \textasciigrave\textasciigrave\textasciigrave. Chunks are supplied with optional arguments within the \textit{\{\}} brackets. A non-exhaustive list of **chunk options**:

Option | Function
---- | -------------
name | A name for the chunk, these have to be individual
eval | `TRUE` if code in chunk should be executed
include | `TRUE` if chunk should be run but not included in final document
echo | `TRUE` if code will be displayed
tidy | `TRUE` if code should be displayed according to `tidy` formatting rules
cache | `TRUE` if cache for chunk is to be stored for faster compilation on repeated knitting

## Functionality
### In-Line Code

In-line code can be used to display code output alongside plain text. This is done as follows:

```{r inline, tidy = FALSE}
# establish and R object to reference in-line
Answer <- 42 
```

Then, you write \textasciigrave$r$ $Answer$\textasciigrave \hspace{.04cm} and receive the output `r Answer`.\newline 

This is especially useful when wanting to generate reports that may be re-evaluated using different analyses or novel data sets as results may change.

### Figures and `Rmarkdown`

*Using* `Rmarkdown` *for data visualisation* is **extremely useful**:
\vspace{.2cm}
\begin{itemize}
\pause
\item Figures are generated on each compilation (unless cache is stored and code in chunk that produces the figure hasn't changed) thus always being up-to-date with the underlying data
\vspace{.1cm}
\item Subfigures are established automatically as long as subcaptions are supplied
\vspace{.1cm}
\item Figures are registered automatically for the table of figures
\vspace{.1cm}
\item Figure placement on the page is fixed (no dragging around as in MS Word)
\vspace{.1cm}
\item No scaling artefacts
\vspace{.1cm}
\item \textbf{Entire workflow can happen in just one document!}
\end{itemize}

\vspace{.2cm}
\pause

`Rmarkdown` **can handle all plotting outputs** of `R` coding including `ggplot2` plots.

# Basic `Rmarkdown`
### Setting Code Up
Basic options for coding chunks and kitting options should be defined at the beginning of the `Rmarkdown` document:

\scriptsize
```{r setupshow, eval=FALSE, tidy = FALSE}
# R options
options(scipen=1, digits=4) # output options such as rounding digits

# Knitting options
knitr::opts_chunk$set(echo = TRUE, # show code
                      message = FALSE, # don't display code messages
                      warning = FALSE, # don't display code warnings 
                      cache = TRUE, # establish chaches
                      cache.lazy=TRUE, # don't check caches on re-run
                      tidy=TRUE, # format tidy by default
                      tidy.opts=list(width.cutoff=50), # width cutoff
                      fig.height=8, digits=4 # default figures sizes
                      )
```

\normalsize
This chunk should usually be set to `include = FALSE` to run the code but not show any of it.

### Text Formatting
Basic `Rmarkdown` formatting can be used to generate more lively text:

\footnotesize
Formatting | Result
-----|-----
Plain text | Plain text
\hspace{.01cm} \*italics\* and \_italics\_ | *italics* and _italics_
\hspace{.01cm} \*\*bold\*\* and \_\_bold\_\_ | **bold** and __bold__
superscript\^2\^ | superscript^2^
$\sim\sim$strikethrough$\sim\sim$ | ~~strikethrough~~

\normalsize

A line break in `Rmakrdown` can be achieved by ending a line with two spaces. \newline

\pause
More advanced text formatting is possible via the \LaTeX \hspace{.04cm} engine. An exhaustive guides for \LaTeX \hspace{.04cm} formatting can be found here (this covers almost all matters \LaTeX \hspace{.04cm}: \url{http://www.rpi.edu/dept/arc/docs/latex/latex-intro.pdf}.

<!-- ### Lists -->

<!-- \begincols -->
<!--   \begincol{.5\linewidth} -->
<!--   The `Rmarkdown` way \newline  -->
<!--   \endcol -->
<!--   \begincol{.5\linewidth} -->
<!--   The \LaTeX way \newline  -->
<!--   \endcol -->
<!-- \endcols -->


### Text Tables

`Rmarkdown` is capable of generating tables through plain text inputs:\newline 

\raggedright
\texttt{Table Header | Second Header} \linebreak
\texttt{------------- | -------------}  \linebreak
\texttt{Table Cell | Cell 2} \linebreak
\texttt{Cell 3 | Cell 4} \newline

\raggedright
turns into:
  
Table Header | Second Header
------------- | -------------
Table Cell | Cell 2
Cell 3 | Cell 4


### Figures
Figures which aren't generated via `R` but present as files on the hard drive have to be included into the `Rmarkdown` document the \LaTeX \hspace{.04cm} way. \newline 

\pause
This is done using the \textbf{includegraphics} command in the \textbf{graphicx} package for \LaTeX. The output of this command is sensitive to other environment specifiers such as floats (\textit{centering, raggedright, etc.}). \newline

\pause
A \LaTeX \hspace{.04cm} \textbf{figure environment} is useful for *including figures* which have been included this way *into the table of figures and make them available for cross-referencing*.

# Advanced `Rmarkdown`
### `kable` Tables I 

The `kable` functionality in `Rmarkdown` enables us to generate appealing \LaTeX styled tables directly via `R` coding: 

\scriptsize
```{r kable, results = 'asis', eval = FALSE}
library(kableExtra)
tabout <- as.data.frame(matrix(c(1:20), nrow = 5))
rownames(tabout) <- c("R1", "R2", "R3", "R4", "R5")
colnames(tabout) <- c("C1", "C2", "C3", "C4")
kab <- kable(tabout, 
            booktabs = TRUE,
            caption = paste("\\textbf{","Short Description", " -} ","Long Description", sep=""),
            caption.short = "Short Description",
            escape = FALSE,
            format = 'latex')
kab <- add_header_above(kable_input = kab, header = c(" " = 1, "First two columns" = 2, "Last two columns" = 2), bold = TRUE, align = 'c')
kab <- kable_styling(kab, latex_options = c("hold_position"))
print(kab)
```
\normalsize
Code chunks which generate `kable` tables need an additional argument: `results = 'asis'`.

### `kable` Tables II

```{r kableshow, ref.label='kable', results = 'asis', echo = FALSE}
```

\vspace{.5cm}

It is obvious how this way of table generation **enables reproducible research** as this tables can be directly generated from data and analyses outputs.

### `R` Functions
Complex data analysis and data visualisation often invokes **user-generated `R` functions**. These should be *presented alongside all other results and data*.\newline

\pause

`Rmarkdown` enables this in a three-step process:  
\vspace{.1cm}
1. **Write Function** in a chunk with `echo` set to `FALSE` (preferably in preamble)  
\vspace{.1cm}
2. **Call Function** into action in another chunk where it is needed  
\vspace{.1cm}
3. **Present Function** using another chunk which is set to `eval = FALSE` by calling the chunk option `ref.label='name'`, where `name` represent the name of the chunk established in step 1

# Manuscripts
### Cross-Referencing
Cross-referencing is vital for manuscripts writing and easy to implement in `Rmarkdown` when run through \LaTeX:
\vspace{-.3cm}
\begincols
  \begincol{.5\linewidth}
  \pause
  
  **Sections**
  \begin{itemize}
  \item Index section by generating a label via the \textit{label} command.
  \item Call section index via the \textit{ref} command.
  \end{itemize}
  \pause
  **Figures**
  \begin{itemize}
  \item Index figure via the \textit{label} command in \LaTeX \hspace{.04cm} environment or chunk option.
  \item Reference using the \textit{ref} command.
  \end{itemize}
  \endcol
  \begincol{.5\linewidth}
  \pause
  **Tables**
  \begin{itemize}
  \item Index table via the \textit{label} command in \LaTeX \hspace{.04cm} environment or chunk option.
  \item Reference using the \textit{ref} command.
  \end{itemize}
  \pause
  **Chunks**
  \begin{itemize}
  \item Generate labelling automation using the \textit{floatrow} package for \LaTeX.
  \item Reference using the \textit{ref} command and name of the chunk in question.
  \end{itemize}
  \endcol
\endcols

### Bibliographies and Glossaries
Including a bibliography or a glossary is a necessity in manuscript writing:
\vspace{-.3cm}

\footnotesize
\begincols
  \begincol{.5\linewidth}  
  \pause
  **Bibliography**
  \begin{itemize}
  \item Install JabRef
  \item Generate .bib bibliography file
  \item In YAML header or auxiliary .tex file included in YAML header, load \LaTeX \hspace{.04cm} package \textit{natbib}
  \item Reference using the \textit{cite} command
  \item Establish bibliography suing the \textit{bibliography command}
  \end{itemize}
  \endcol
  \begincol{.5\linewidth}  
  \pause
  **Glossary**
  \begin{itemize}
  \item Install Perl
  \item In YAML header or auxiliary .tex file included in YAML header, load \LaTeX \hspace{.04cm} package \textit{glossaries}
  \item Establish glossary entries in .tex file or preamble to document file
  \item Establish glossary at desired location in document using the \textit{printglossary} command
  \item Use glossary commands in text
  \item Compiling requires an extra step through your \LaTeX \hspace{.04cm} compiler for glossary to show up
  \end{itemize}
  \endcol
\endcols

# Presentations
### Beamer
`Beamer` is the \LaTeX \hspace{.04cm} way of generating presentation slides and comes with a few quirks:
\vspace{.1cm}
\pause
\begin{itemize}
\item Strong limitations on layout (it is best to chose from pre-established styles)
\vspace{.1cm}
\item No animations possible
\vspace{.1cm}
\item Stiff and unwieldy positioning of items
\end{itemize}
\pause
\vspace{.1cm}
These limitations often lead to *headaches when generating the presentations* but **much cleaner looking presentations** with **consistent formatting**.

\pause
\vspace{.3cm}
\LaTeX \hspace{.04cm} column environment (used in `Beamer` presentations) do not allow for the evaluation of `Rmarkdown` syntax. 
\vspace{.1cm}
This can be circumvented as by defining novel commands for the column environment instead of calling the default \LaTeX\hspace{.04cm} commands.


