---
title: "Coding Practices"
subtitle: "Life with R"
author: "Erik Kusch"
date: "25/03/2020"
fontsize: 10pt
output: 
  beamer_presentation: 
    keep_tex: true # keep latex file that is generated during compilation
    toc: false # this is added through a later command
    slide_level: 3 # at how many pound signs (#) to assume slide title level
    includes:
      in_header: Style.tex
classoption: c # change to `handout` to ignore breaks
---
  
```{r setup, include=FALSE}
options(scipen=1, digits=4)
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, cache = TRUE, cache.lazy=TRUE, tidy.opts=list(width.cutoff=50),tidy=TRUE, fig.height=8, digits=4)
```

<!--- ####### TOC (use this if you want ## headers to be included in the toc)--------------------
---------------------------------------  --->
  
\tableofcontents

<!--- ####### DOCUMENT --------------------
---------------------------------------  --->

# Quality of Life Improvements
## Coding in `R`
### Objects

\begin{center}
Assign \textbf{recognizable}, \textbf{concise} names to your objects!
\end{center}

\pause
\vspace{.4cm}

Let's start with the object `MyData` - this can be anything. Let's assume it is a raster of NDVI values during the year 1982. How do we come up with a better name?

\begincols[T]
\begincol{.5\linewidth}
\pause
  **Specificity** in your naming helps you keep track of your data in your code:  
  - `NDVI1982`  

\endcol
\begincol{.5\linewidth}
\pause
  **Classes** of objects determine how they get handled. Add them as a suffix:  
  - `NDVI1982_ras`  
  
\endcol
\endcols

\vspace{.4cm}
\pause
The same goes for **loop indicators**! Avoid single-letter indicators and instead use something like `Iter_Years` - an indicator *iterating* (`Iter`) over years.

\vspace{.5cm}
\pause
\raggedleft Also, always use $<-$ for assigning objects!

### Quality of Life Improvements

Use **spaces** to improve readability of your code:
\footnotesize
```{r Spaces, eval = FALSE}
paste0("Bad","Spacing") # bad
paste0("Good", "Spacing") # good
```
\normalsize
\vspace{.3cm}

\pause 
Insert **line breaks** to improve workflow. I like to break multiple lines of one function call into logical blocks. For example, in `ggplot2`, the first line holds data, the second contains aesthetics, the third adds a layer, and so on.

\vspace{.3cm}
\pause 
To avoid side-scrolling, use the automatic **soft-wrap** function in `RStudio`.

\vspace{.3cm}
\pause 
Be **consistent** in your coding style. Found a `tidyverse` solution to your problem online but don't use `tidyverse` syntax in the rest of your document? Adapt the solution to be in the same style as the rest of your coding!


## Documenting Your Work
### Comments

\begin{center}
Don't be overly proud of your coding skills. Comment everything!
\end{center}

\vspace{.3cm}

  \pause
  \textbf{Bad comments:}  
    - ambiguous  
    - sparse  
    
\footnotesize
```{r CommentsB, eval = FALSE}
# Data Manipulation
```
\normalsize

  \pause
  \textbf{Good comments:}  
    - specify (what is the code doing)  
    - justify (why is it being done)  
  
  \footnotesize
```{r CommentsG, eval = FALSE}
# Z-Score calculation for comparability
```  
\normalsize

\vspace{.3cm}
\pause
\begin{tcolorbox}[colback=green!5,colframe=green!40!black,title=]
\centering Write comments as though you were \textbf{coding for someone else}!
\end{tcolorbox}


### `RMarkdown`
Using `Rmarkdown` for your research comes with a multitude of advantages:  \vspace{.3cm}

1. Entire **workflow in one program** (`RStudio`)  
\vspace{.1cm}
2. Great capabilities to **present code**  
\vspace{.1cm}
3. **Research** and reports **reproducible** at the click of **one button**  
\vspace{.1cm}
4. **Combines** `R` functionality and \LaTeX \hspace{.04cm} formatting (if desired)  
\vspace{.1cm}
5. **Consistent formatting**  
\vspace{.1cm}
6. **Clear presentation of code**  
\vspace{.1cm}
7. **Dynamic documents** (you can generate various output document types)  
\vspace{.1cm}
8. Applicable for **almost all document types** you may desire as an output (e.g. manuscripts, presentations, posters, etc.)  


# Code Structure
### Header
The **Head** is used as an information statement at the top of your code document that informs the user of the contents, author, and (sometimes) date of the last edit on said document:
\vspace{.2cm}
\footnotesize
```{r Head, eval = FALSE, tidy = FALSE}
# ################################################# #
# PROJECT: Project Title
# CONTENTS: What the code in your file is used for
# AUTHOR: Who worked on it
# EDIT: When the last edit was made
# ################################################# #
```
\normalsize
\pause
\vspace{.2cm}
Personally, I don't use the edit argument in my heads. Instead, I rely on *version control*.

### Preamble
The **Preamble** is where you set up the most important parameters/guidelines for your coding script. Personally, I *highly* recommend to make your first line in the preamble read `rm(list=ls())`.  

This is also where you load **packages**. I recommend doing so as follows:
\footnotesize
```{r Packages1, eval = FALSE}
install.load.package <- function(x) {
  if(!require(x, character.only = TRUE))
    install.packages(x, repos='http://cran.us.r-project.org')
  require(x, character.only = TRUE)}
```
\normalsize
This functions *checks* whether the package is already installed. If it is, it will be *loaded*. If it is not, it will be *installed* from `CRAN` and subsequently loaded.  

You use it for multiple packages like this:
```{r Packages2, eval = FALSE}
package_vec <- c("ggplot2", "foreach", "doParallel")
sapply(package_vec, install.load.package)
```

### Sections

**Sections** are a powerful tool in `RStudio`! 
\vspace{.2cm}

\pause
\begincols[T]
\begincol{.5\linewidth}
  They keep your code:  
- *structured*  
- *clear*  
- *easy-to-use*
\endcol
\begincol{.5\linewidth}
\pause
  They do so by:  
- *collapsing* code  
- adding a *table of contents* 
\endcol
\endcols

\centering \includegraphics[width=0.8\linewidth]{Figures/Sections.png}

# Sharing Work
## Reproducibility
### Reproducibility
**Reproducibility** is **paramount** in science.
\vspace{.2cm}

\pause
\begincols[T]
\begincol{.5\linewidth}
  **Working directories**  
  It is common to set a base directory at the beginning of your code and refer to sub-directories from there. We can *do better*:  
  - `getwd()` in a `project` will yield the directory of the `project` file  
  - the package `here` offers more functions to never have to set a directory path by hand  

\endcol
\begincol{.5\linewidth}
\pause
  **Randomness**  
  Random processes (e.g. `sample()`, `rnorm()`, etc.) in `R` are always *pseudo-random* (their calculation is based on your computer time). We can make sure the result of a random process is always the same between researchers by using:  
  - `set.seed()`  
  
\endcol
\endcols

## Functions \& Sourcing
### Functions \& Sourcing
*Big projects* lend themselves well to a **multi-coment workflow**:
\vspace{.2cm}

\pause
\begincols[T]
\begincol{.5\linewidth}
  **Functions**  
  User-defined functions are a great way of *soft-coding* your analyses to repeat them for different input parameters and I highly recommend doing so all the time. Make sure your functions are:  
  - internally *consistent*  
  - *well-documented*  
  - easy to *understand*  
  
\endcol
\begincol{.5\linewidth}
\pause
  **Sourcing**  
  Splitting your code into multiple documents is a great idea to ensure your project remains *structured* and *easy to handle*. You can use the `source()` function to combine those documents. Keep in mind that you should:  
  - use *sensible file names*  
  - *sourced* functions still need to be *called*
  
\endcol
\endcols

# User-Friendliness
## Progress Bar
### Progress Bar
A **progress bar** is a great way of keeping you updated on how your code is progressing so far. It is especially useful when your code involves *loops*:
\vspace{.2cm}
\footnotesize
```{r Prog, eval = FALSE, tidy = FALSE}
Data_vec <- 1:100 # a vector on integers from 1 to 100
# creating progress bar
ProgBar <- txtProgressBar(min = 0, max = length(Data_vec), 
                          style = 3)
# looping over contents of Data_vec
for(Iter_ProgBar in 1:length(Data_vec)){
  setTxtProgressBar(ProgBar, Iter_ProgBar) # updating ProgBar
} # end of Data_vec loop
```
\centering \includegraphics[width=1\linewidth]{Figures/ProgBar.png}

## Estimators
### Estimators
**Estimators** are a great way to know when to come back to your computer or server and check up on your data, code, and results. They are most useful in *loop* based approaches as they need a baseline for the estimation:
\vspace{.2cm}
\tiny
```{r Est, tidy = FALSE}
Data_vec <- 1:100 # a vector on integers from 1 to 100
T_Begin <- Sys.time() # record time
# looping over contents of Data_vec
for(Iter_Est in 1:length(Data_vec)){
  Sys.sleep(.1) # pause for .1 seconds
  # estimator produced on first iteration
  if(Iter_Est == 1){
    T_End <- Sys.time() # record time
    Duration <- as.numeric(T_End)-as.numeric(T_Begin) # time difference
    print(paste("Estimated time to finish:", 
                as.POSIXlt(T_Begin + Duration*length(Data_vec), 
                           tz = Sys.timezone(location=FALSE))
                ))
  } # end of estimator check
} # end of Data_vec loop
```
\vspace{.2cm}
\raggedleft Yes, I did work on this presentation past midnight.

## Parallel Processing
### Parallel Processing
Got a big data approach where you carry out the same kind of analysis for many different species or data sets? It's taking forever to complete? **Parallel processing** has your back! By default, `R` only uses one of the cores in your computer or server for processing code. Use this, to make use of all the computational power:
\footnotesize
```{r Par, eval = FALSE, tidy = FALSE}
library(doParallel) # for registering clusters
library(foreach) # for parallel processing
Cores <- detectCores() # identify the number of cores in your machine
cl <- makeCluster(Cores) # create virtual cluster
registerDoParallel(cl) # register cluster of cores
# parallel processing
foreach(Iter_Par = 1:length(Data_vec)) %dopar% {
  # your function here
} # end of parallel processing
stopCluster(cl) # stop cluster

```
